---
title: "Posterior Sampling Time Analysis -- $m=64$ and $T=500$"
author: Yifan CHENG $\quad$ [\textcolor{blue}{y.cheng@u.nus.edu}](mailto:y.cheng@u.nus.edu)  
output:
  pdf_document: default
header-includes: 
 - \usepackage{bm}
 - \usepackage{xcolor}
---

# Simulation Set-Up and Overall Model Fitting Time
When fitting each of our 4 models (`fullGPfixedL`, `NNGPblockFixedL`, `NNGPsequenFixedL`, and `NNGPsequenVaryLj`) with $m=8^2=64$ and $T=500$, we specified both `equalTimeDist = TRUE` and `equalTimeDist = FALSE`. We thus have $4\times 2 = 8$ model settings. 
```{r eval=F}
setwd(paste(projDirec, "simu/appen/equalTimeDistAccelerationVerification/m64T500", sep = "/"))
load("regFixedL30simuT500M64Iter30000.RData")
regFixedL.simu.fast <- regFixedL.simu
load("regFixedL30simuT500M64Iter30000specifyEqualTimeDistF.RData")
load("regFixedL30simuBlockT500M64Iter30000.RData")
regFixedL.simu.block.fast <- regFixedL.simu.block
load("regFixedL30simuBlockT500M64Iter30000specifyEqualTimeDistF.RData")
load("regFixedL30simuSequenT500M64Iter30000nostorealphaweights.RData")
regFixedL.simu.sequen.fast <- regFixedL.simu.sequen
load("regFixedL30simuSequenT500M64Iter30000specifyEqualTimeDistFnostorealphaweights.RData")
load("regVaryLjsimuSequenT500M64Iter30000nostorealphaweight.RData")
regVaryLj.simu.sequen.fast <- regVaryLj.simu.sequen
load("regVaryLjsimuSequenT500M64Iter30000specifyEqualTimeDistFnostorealphaweight.RData")
##all of the 8 matrices below are 5000 x 10 (10000 / 2 = 5000 kept post-burn-in MCMC iterations, 7 + 3 = 10 key parameters)## 
GibbsStepTimeFixedLfullGP.fast <- regFixedL.simu.fast$GibbsStepTime         # equalTimeDist = TRUE
GibbsStepTimeFixedLblock.fast <- regFixedL.simu.block.fast$GibbsStepTime    # equalTimeDist = TRUE
GibbsStepTimeFixedLsequen.fast <- regFixedL.simu.sequen.fast$GibbsStepTime  # equalTimeDist = TRUE
GibbsStepTimeVaryLjSequen.fast <- regVaryLj.simu.sequen.fast$GibbsStepTime  # equalTimeDist = TRUE
GibbsStepTimeFixedLfullGP <- regFixedL.simu$GibbsStepTime                   # equalTimeDist = FALSE
GibbsStepTimeFixedLblock <- regFixedL.simu.block$GibbsStepTime              # equalTimeDist = FALSE
GibbsStepTimeFixedLsequen <- regFixedL.simu.sequen$GibbsStepTime            # equalTimeDist = FALSE
GibbsStepTimeVaryLjSequen <- regVaryLj.simu.sequen$GibbsStepTime            # equalTimeDist = FALSE
save(GibbsStepTimeFixedLfullGP, file = "GibbsStepTimeFixedLfullGP.RData")
save(GibbsStepTimeFixedLblock, file = "GibbsStepTimeFixedLblock.RData")
save(GibbsStepTimeFixedLsequen, file = "GibbsStepTimeFixedLsequen.RData")
save(GibbsStepTimeVaryLjSequen, file = "GibbsStepTimeVaryLjSequen.RData")
save(GibbsStepTimeFixedLfullGP.fast, file = "GibbsStepTimeFixedLfullGPfast.RData")
save(GibbsStepTimeFixedLblock.fast, file = "GibbsStepTimeFixedLblockFast.RData")
save(GibbsStepTimeFixedLsequen.fast, file = "GibbsStepTimeFixedLsequenFast.RData")
save(GibbsStepTimeVaryLjSequen.fast, file = "GibbsStepTimeVaryLjSequenFast.RData")
```
Since $m=64$ is quite small, we do not expect any significant differences in the posterior sampling time between our 4 models with the same `equalTimeDist` input. Since $T=500$ is quite large, what we have discussed in Appendix B of our manuscript suggest that the 3 temporal parameters' posterior sampling steps, especially the step for $\bm{\eta}_t$'s, can be markedly accelerated by adopting our approaches and specifying `equalTimeDist = TRUE` for each of our 4 models. 

The results we have obtained corroborate these well. With `equalTimeDist = TRUE` specified, the overall model fitting time for our 4 models, i.e., `fullGPfixedL`, `NNGPblockFixedL`, `NNGPsequenFixedL`, and `NNGPsequenVaryLj`, are 7.04 hours, 7.03 hours, 7.04 hours, and 6.31 hours, respectively.^[It is worth noting that we utilized a version of our package with the extra step-wise time recording code portions to fit the above models. This package version is significantly slower than our original package version presented in the GitHub repository.] If we do not take advantage of our tactics for evenly dispersed time points presented in Appendix B of our manuscript by specifying `equalTimeDist = FALSE` instead, we will need 29.03 days, and 28.74 days to fit the same 4 models using the same computation resources. 

# Portions of Recorded Gibbs Sampler Time for 10 Key Parameters
We first display the first 50 kept post-burn-in MCMC iterations' posterior sampling time (in milliseconds) for 10 key Gibbs sampler steps (corresponding to $\psi$, $\Upsilon$, $\bm{\eta}_t$'s, $z_{jl}(\bm{s}_{i,o})$'s or $u_j(\bm{s}_{i,o})$'s, $\xi_j(\bm{s}_{i,o})$'s, $\theta_{jl_j}$'s, $\delta_{1:k}$, $\rho$, $\kappa$, and $\alpha_{jl_j}(\bm{s}_{i,o})$'s) for our 8 model settings.
```{r echo = -(1:2)}
rm(list=ls())
projDirec <- "C:/Users/annie/OneDrive - National University of Singapore/Documents/PhD/research/first paper/spatempBFA"
wd <- paste(projDirec, "simu/appen/equalTimeDistAccelerationVerification/m64T500", sep = "/")
setwd(wd)
load("GibbsStepTimeFixedLfullGP.RData")
load("GibbsStepTimeFixedLfullGPfast.RData")
load("GibbsStepTimeFixedLblock.RData")
load("GibbsStepTimeFixedLblockFast.RData")
load("GibbsStepTimeFixedLsequen.RData")
load("GibbsStepTimeFixedLsequenFast.RData")
load("GibbsStepTimeVaryLjSequen.RData")
load("GibbsStepTimeVaryLjSequenFast.RData")
head(GibbsStepTimeFixedLfullGP, 50)      # equalTimeDist = FALSE
head(GibbsStepTimeFixedLfullGP.fast, 50) # equalTimeDist = TRUE
head(GibbsStepTimeFixedLblock, 50)       # equalTimeDist = FALSE
head(GibbsStepTimeFixedLblock.fast, 50)  # equalTimeDist = TRUE
head(GibbsStepTimeFixedLsequen, 50)      # equalTimeDist = FALSE
head(GibbsStepTimeFixedLsequen.fast, 50) # equalTimeDist = TRUE
head(GibbsStepTimeVaryLjSequen, 50)      # equalTimeDist = FALSE
head(GibbsStepTimeVaryLjSequen.fast, 50) # equalTimeDist = TRUE
```   

# Posterior Sampling Time Summary Statistics for the 3 Temporal Parameters
We then present vital posterior sampling time summary statistics for the 3 temporal parameters $\psi$, $\Upsilon$, and $\bm{\eta}_t$'s to showcase the manifest computational acceleration brought about by our tactics for evenly dispersed time points presented in Appendix B.
```{r}
fullGPfixedL.slow <- apply(GibbsStepTimeFixedLfullGP[,8:10], 2, summary)      # equalTimeDist = FALSE
fullGPfixedL.fast <- apply(GibbsStepTimeFixedLfullGP.fast[,8:10], 2, summary) # equalTimeDist = TRUE
NNGPblockFixedL.slow <- apply(GibbsStepTimeFixedLblock[,8:10], 2, summary)       # equalTimeDist = FALSE
NNGPblockFixedL.fast <- apply(GibbsStepTimeFixedLblock.fast[,8:10], 2, summary)  # equalTimeDist = TRUE
NNGPsequenFixedL.slow <- apply(GibbsStepTimeFixedLsequen[,8:10], 2, summary)      # equalTimeDist = FALSE
NNGPsequenFixedL.fast <- apply(GibbsStepTimeFixedLsequen.fast[,8:10], 2, summary) # equalTimeDist = TRUE
NNGPsequenVaryLj.slow <- apply(GibbsStepTimeVaryLjSequen[,8:10], 2, summary)      # equalTimeDist = FALSE
NNGPsequenVaryLj.fast <- apply(GibbsStepTimeVaryLjSequen.fast[,8:10], 2, summary) # equalTimeDist = TRUE
fullGPfixedLsummary <- data.frame(eta = fullGPfixedL.slow[,1], eta.fast = fullGPfixedL.fast[,1],
                                  upsilon = fullGPfixedL.slow[,2], upsilon.fast = fullGPfixedL.fast[,2],
                                  psi = fullGPfixedL.slow[,3], psi.fast = fullGPfixedL.fast[,3])
NNGPblockFixedLsummary <- data.frame(eta = NNGPblockFixedL.slow[,1], eta.fast = NNGPblockFixedL.fast[,1],
                                  upsilon = NNGPblockFixedL.slow[,2], upsilon.fast = NNGPblockFixedL.fast[,2],
                                  psi = NNGPblockFixedL.slow[,3], psi.fast = NNGPblockFixedL.fast[,3])
NNGPsequenFixedLsummary <- data.frame(eta = NNGPsequenFixedL.slow[,1], eta.fast = NNGPsequenFixedL.fast[,1],
                                      upsilon = NNGPsequenFixedL.slow[,2], upsilon.fast = NNGPsequenFixedL.fast[,2],
                                      psi = NNGPsequenFixedL.slow[,3], psi.fast = NNGPsequenFixedL.fast[,3])
NNGPsequenVaryLjsummary <- data.frame(eta = NNGPsequenVaryLj.slow[,1], eta.fast = NNGPsequenVaryLj.fast[,1],
                                      upsilon = NNGPsequenVaryLj.slow[,2], upsilon.fast = NNGPsequenVaryLj.fast[,2],
                                      psi = NNGPsequenVaryLj.slow[,3], psi.fast = NNGPsequenVaryLj.fast[,3])
fullGPfixedLsummary
NNGPblockFixedLsummary 
NNGPsequenFixedLsummary
NNGPsequenVaryLjsummary
```
The results correspond well to what we have discussed in Appendix B of our manuscript. **For each of our 4 models, specifying `equalTimeDist = TRUE` instead of `equalTimeDist = FALSE` on equispaced time points markedly accelerates posterior Gibbs sampler steps of $\bm{\eta}_t$'s, $\psi$, and $\Upsilon$.**

We finally calculate standard deviations for the 3 temporal-related parameters' posterior sampling time across all kept post-burn-in MCMC iterations.
```{r}
round(apply(GibbsStepTimeFixedLfullGP[,8:10], 2, sd), 5)
round(apply(GibbsStepTimeFixedLfullGP.fast[,8:10], 2, sd), 5)
round(apply(GibbsStepTimeFixedLblock[,8:10], 2, sd), 5)
round(apply(GibbsStepTimeFixedLblock.fast[,8:10], 2, sd), 5)
round(apply(GibbsStepTimeFixedLsequen[,8:10], 2, sd), 5)
round(apply(GibbsStepTimeFixedLsequen.fast[,8:10], 2, sd), 5)
round(apply(GibbsStepTimeVaryLjSequen[,8:10], 2, sd), 5)
round(apply(GibbsStepTimeVaryLjSequen.fast[,8:10], 2, sd), 5)
```
