---
title: "Key Parameters' Posterior Sampling Time Analysis"
author: Yifan CHENG $\quad$ [\textcolor{blue}{y.cheng@u.nus.edu}](mailto:y.cheng@u.nus.edu)  
output:
  pdf_document: default
header-includes: 
 - \usepackage{bm}
 - \usepackage{xcolor}
---

# Portions of Recorded Gibbs Sampler Time for 10 Key Parameters
We first display the first 50 kept post-burn-in MCMC iterations' posterior sampling time (in milliseconds) for 10 key Gibbs sampler steps corresponding to our 4 models, i.e., `fullGPfixedL`, `NNGPblockFixedL`, `NNGPsequenFixedL`, and `NNGPsequenVaryLj`.
```{r echo = -(1:2)}
rm(list=ls())
projDirec <- "C:/Users/annie/OneDrive - National University of Singapore/Documents/PhD/research/first paper/spatempBFA"
wd <- paste(projDirec, "simu/mainScalabilityVerificationSimu/m3600T100K5", sep = "/")
setwd(wd)
load("GibbsStepTimeFixedLfullGP.RData")
load("GibbsStepTimeFixedLblock.RData")
load("GibbsStepTimeFixedLsequen.RData")
load("GibbsStepTimeVaryLjSequen.RData")
head(GibbsStepTimeFixedLfullGP, 50)
head(GibbsStepTimeFixedLblock, 50)
head(GibbsStepTimeFixedLsequen, 50)
head(GibbsStepTimeVaryLjSequen, 50)
```   
As expected, there aren't any significant differences between our 4 models regarding posterior sampling time for the 3 temporal parameters $\psi$, $\Upsilon$, and $\bm{\eta}_t$'s.

# Posterior Sampling Time Summary Statistics
We then present vital posterior sampling time summary statistics for the 7 spatial-related parameters ($z_{jl}(\bm{s}_{i,o})$'s or $u_j(\bm{s}_{i,o})$'s, $\xi_j(\bm{s}_{i,o})$'s, $\theta_{jl_j}$'s, $\delta_{1:k}$, $\rho$, $\kappa$, and $\alpha_{jl_j}(\bm{s}_{i,o})$'s) to showcase the manifest scalability improvements brought about by our 3 novelties, i.e., slice sampling, spatial NNGP, and sequential updates.
```{r}
apply(GibbsStepTimeFixedLfullGP[,1:7], 2, summary)
apply(GibbsStepTimeFixedLblock[,1:7], 2, summary)
apply(GibbsStepTimeFixedLsequen[,1:7], 2, summary)
apply(GibbsStepTimeVaryLjSequen[,1:7], 2, summary)
```
The results correspond well to what we have deduced in Appendix H of our manuscript. 

- Compared to their `fullGPfixedL` counterparts, `NNGPblockFixedL`'s Gibbs sampler steps corresponding to $\rho$ and $\kappa$ are evidently accelerated by our **spatial NNGP prior**;
- The only Gibbs sampler step time that should clearly differ between `NNGPblockFixedL` and `NNGPsequenFixedL` is the step updating all $\alpha_{jl}(\bm{s}_{i,o})$'s, which result from whether we adopt our **sequential updating method** or not. Since $m=3600$ here is big, `NNGPsequenFixedL` is more than 20 times faster than `NNGPblockFixedL` for the posterior sampling step corresponding to $\alpha_{jl}(\bm{s}_{i,o})$'s;
- Thanks to our **slice sampling approach**, `NNGPsequenVaryLj`'s Gibbs sampler steps for $u_j(\bm{s}_{i,o})$'s, $\xi_j(\bm{s}_{i,o})$'s, $\rho$, and $\kappa$ are significantly faster than `NNGPsequenFixedL`'s Gibbs sampler steps for $z_{jl}(\bm{s}_{i,o})$'s, $\xi_j(\bm{s}_{i,o})$'s, $\rho$, and $\kappa$. It turns out that `NNGPsequenVaryLj`'s Gibbs sampler step for $\alpha_{jl_j}(\bm{s}_{i,o})$'s is slower than its `NNGPsequenFixedL` counterpart, indicating that inefficiencies caused by case discussion, calculating all required upper or lower bounds, and rejection sampling outweigh acceleration brought about by slice sampling's ensured non-increasing posterior samples for $L_j$'s through the MCMC iterations.

We finally calculate standard deviations for the 7 spatial-related parameters' posterior sampling time across all kept post-burn-in MCMC iterations.
```{r}
round(apply(GibbsStepTimeFixedLfullGP[,1:7], 2, sd), 5)
round(apply(GibbsStepTimeFixedLblock[,1:7], 2, sd), 5)
round(apply(GibbsStepTimeFixedLsequen[,1:7], 2, sd), 5)
round(apply(GibbsStepTimeVaryLjSequen[,1:7], 2, sd), 5)
```

# Data Generation and Model Fitting
```{r eval = FALSE}
rm(list=ls())
library(mvtnorm)
library(fields)
library(spatempBFA)
library(coda)
K <- 5
O <- 1
L <- 50
M <- 3600
LjVec <- rep(min(50, M), K)
sqrootM <- 60
Nu <- 100
Time <- 1:Nu
TimeDist <- as.matrix(dist(Time))
APsi = 0.1; BPsi = 4.5
set.seed(29)
### 1) actual sigma^2(i,o) (for i=1,2,...,M and o=1) values
sigma2 <- 0.01
### 2) actual psi value
psi <- 2.3
### 3) actual kappa value
kappa <- 0.7
### 4) actual Upsilon(K\times K)
tempMat <- matrix(runif(K*K,0,1),K,K)
Upsilon <- t(tempMat)%*%tempMat
### 5) actual rho value
rho <- 0.8
D <- rdist(expand.grid(1:sqrootM, 1:sqrootM))
Frho <- exp(-rho*D)
### 6) actual Eta (c(Eta_1,...,Eta_T)) (vec of length Nu*K)
Hpsi <- exp(-psi*TimeDist)
Eta <- rmvnorm(1, mean=rep(0, Nu*K), sigma=kronecker(Hpsi, Upsilon))
## Y~0 (P=0) so no need to sample Beta; all familyInd=0 (all normal) so no need to sample Y
maxL <- 10
LStarJ <- sample(maxL, size=K, replace=T)
### 7) actual alpha
Alpha <- list()
for(j in 1:K) {
  Alpha[[j]] <-  t(rmvnorm(LStarJ[j], mean=rep(0,M*O), sigma=kappa*Frho))
  #every list index an M by L_j matrix 
}
w <- list()
for(j in 1:K){
  w[[j]] <- pnorm(Alpha[[j]])
  Lj <- LStarJ[j]
  w[[j]][,Lj] <- rep(1, M)
  temp <- rep(1, M)
  for(l in 1:Lj){
    w[[j]][,l] <- w[[j]][,l]*temp
    if(l<Lj) {temp <- temp * pnorm(Alpha[[j]][,l], lower.tail = FALSE)}
  }
}
### 8) actual Xi
Xi <- matrix(1, M, K)
for(j in 1:K){
  Lj <- LStarJ[j]
  for(i in 1:M){
    Xi[i,j] <- sample(Lj, size=1, prob=w[[j]][i,])
  }
}
### 9) actual Delta
a1=1; a2=10
Delta <- sapply(c(a1,rep(a2,(K-1))), rgamma, n=1, rate=1)
#Tau <- cumprod(Delta)
### 10) actual Theta
Theta <- list()
for(j in 1:K){
  Theta[[j]] <- rnorm(LStarJ[j], 0, sd=sqrt(1/Delta[j])) #vector of length Lj
}
Lambda <- matrix(0, M*O, K)
for(j in 1:K){
  for(i in 1:M){
    Lambda[i,j] = Theta[[j]][Xi[i,j]]
  }
}
Sigma.NuMO <- rnorm(Nu * M * O, sd = sqrt(sigma2))
EtaMat <- matrix(Eta, K, Nu)
meanMat <- Lambda%*%EtaMat #M*O\times Nu
Yobs <- as.vector(meanMat) + Sigma.NuMO
dat <- data.frame(Y = Yobs)
Hypers <- list(Sigma2 = list(A = 1, B = 1), Rho = list(ARho=0.1, BRho=1),
               Kappa = list(SmallUpsilon = O + 1, BigTheta = diag(O)),
               Psi = list(APsi = APsi, BPsi = BPsi),
               Upsilon = list(Zeta = K + 1, Omega = diag(K)))
MCMC <- list(NBurn = 20000, NSims = 10000, NThin = 2, NPilot = 5)
regFixedL.simu <- bfaFixedL(Y ~ 0, data = dat, dist = D, time = Time,  K = K, 
                            starting = NULL, hypers = Hypers, tuning = NULL, mcmc = MCMC,
                            L = L,
                            family = "normal",
                            temporal.structure = "exponential",
                            spatial.structure = "continuous",
                            seed = 29, 
                            gamma.shrinkage = TRUE,
                            include.time = TRUE,
                            include.space = TRUE,
                            clustering = TRUE,
                            seasonPeriod = 1, 
                            equalTimeDist = TRUE,
                            spatApprox = FALSE, 
                            alphaMethod = "block", 
                            h = 15, 
                            storeSpatPredPara = FALSE,
                            storeWeights = FALSE,
                            alphasWeightsToFiles = FALSE) 
save(regFixedL.simu, file="regFixedL50simuT100M3600Iter30000.RData")
Diags <- diagnostics(regFixedL.simu, diags = c("dic", "dinf", "meanIC", "waic"), 
                     keepDeviance = TRUE)
save(Diags, file = "regFixedL50simuT100M3600Iter30000Diags.RData")
Deviance <- as.mcmc(Diags$deviance)
save(Deviance, file = "regFixedL50simuT100M3600Iter30000Deviance.RData")
GibbsStepTimeFixedLfullGP <- regFixedL.simu$GibbsStepTime
save(GibbsStepTimeFixedLfullGP, file = "GibbsStepTimeFixedLfullGP.RData")
regFixedL.simu.block <- bfaFixedL(Y ~ 0, data = dat, dist = D, time = Time,  K = K, 
                                  starting = NULL, hypers = Hypers, tuning = NULL, 
                                  mcmc = MCMC,
                                  L = L,
                                  family = "normal",
                                  temporal.structure = "exponential",
                                  spatial.structure = "continuous",
                                  seed = 29, 
                                  gamma.shrinkage = TRUE,
                                  include.time = TRUE,
                                  include.space = TRUE,
                                  clustering = TRUE,
                                  seasonPeriod = 1, 
                                  equalTimeDist = TRUE,
                                  spatApprox = TRUE, 
                                  alphaMethod = "block", 
                                  h = 15,
                                  storeSpatPredPara = FALSE,
                                  storeWeights = FALSE,
                                  alphasWeightsToFiles = FALSE) 
save(regFixedL.simu.block, file="regFixedL50simuBlockT100M3600Iter30000.RData")
Diags.block <- diagnostics(regFixedL.simu.block, 
                           diags = c("dic", "dinf", "meanIC", "waic"), 
                           keepDeviance = TRUE)
save(Diags.block, file = "regFixedL50simuBlockT100M3600Iter30000Diags.RData")
Deviance.block <- as.mcmc(Diags.block$deviance)
save(Deviance.block, file = "regFixedL50simuBlockT100M3600Iter30000Deviance.RData")
GibbsStepTimeFixedLblock <- regFixedL.simu.block$GibbsStepTime
save(GibbsStepTimeFixedLblock, file = "GibbsStepTimeFixedLblock.RData")
regFixedL.simu.sequen <- bfaFixedL(Y ~ 0, data = dat, dist = D, time = Time,  K = K, 
                                   starting = NULL, hypers = Hypers, tuning = NULL, 
                                   mcmc = MCMC,
                                   L = L,
                                   family = "normal",
                                   temporal.structure = "exponential",
                                   spatial.structure = "continuous",
                                   seed = 29, 
                                   gamma.shrinkage = TRUE,
                                   include.time = TRUE,
                                   include.space = TRUE,
                                   clustering = TRUE,
                                   seasonPeriod = 1, 
                                   equalTimeDist = TRUE,
                                   spatApprox = TRUE, 
                                   alphaMethod = "sequential", 
                                   h = 15, 
                                   storeSpatPredPara = FALSE,
                                   storeWeights = FALSE,
                                   alphasWeightsToFiles = FALSE) 
save(regFixedL.simu.sequen, 
     file="regFixedL50simuSequenT100M3600Iter30000nostorealphaweights.RData")
Diags.sequen <- diagnostics(regFixedL.simu.sequen, 
                            diags = c("dic", "dinf", "meanIC", "waic"), 
                            keepDeviance = TRUE)
save(Diags.sequen, 
     file = "regFixedL50simuSequenT100M3600Iter30000DiagsNostorealphaweights.RData")
Deviance.sequen <- as.mcmc(Diags.sequen$deviance)
save(Deviance.sequen, 
     file = "regFixedL50simuSequenT100M3600Iter30000DevianceNostorealphaweights.RData")
GibbsStepTimeFixedLsequen <- regFixedL.simu.sequen$GibbsStepTime
save(GibbsStepTimeFixedLsequen, file = "GibbsStepTimeFixedLsequen.RData")
regVaryLj.simu.sequen <- bfaVaryingLjs(Y ~ 0, data = dat, dist = D, time = Time,  K = K, 
                                       LjVec = LjVec,
                                       starting = NULL, hypers = Hypers, tuning = NULL, 
                                       mcmc = MCMC,
                                       family = "normal",
                                       temporal.structure = "exponential",
                                       spatial.structure = "continuous",
                                       seed = 29, 
                                       gamma.shrinkage = TRUE,
                                       include.time = TRUE,
                                       include.space = TRUE,
                                       seasonPeriod = 1, 
                                       equalTimeDist = TRUE,
                                       spatApprox = TRUE, 
                                       alphaSequen = TRUE, 
                                       h = 15,
                                       storeSpatPredPara = FALSE, 
                                       storeWeights = FALSE) 
save(regVaryLj.simu.sequen, 
     file="regVaryLjsimuSequenT100M3600Iter30000nostorealphaweight.RData")
Diags.sequenVaryLj <- diagnostics(regVaryLj.simu.sequen, 
                                  diags = c("dic", "dinf", "meanIC", "waic"),
                                  keepDeviance = TRUE)
save(Diags.sequenVaryLj, 
     file = "regVaryLjsimuSequenT100M3600Iter30000DiagsNostorealphaweight.RData")
Deviance.sequenVaryLj <- as.mcmc(Diags.sequenVaryLj$deviance)
save(Deviance.sequenVaryLj, 
     file = "regVaryLjsimuSequenT100M3600Iter30000DevianceNostorealphaweight.RData")
GibbsStepTimeVaryLjSequen <- regVaryLj.simu.sequen$GibbsStepTime
save(GibbsStepTimeVaryLjSequen, 
     file = "GibbsStepTimeVaryLjSequen.RData")
```




     
     


     




