---
title: "Key Parameters' Posterior Sampling Time Analysis"
author: Yifan CHENG $\quad$ [\textcolor{blue}{y.cheng@u.nus.edu}](mailto:y.cheng@u.nus.edu)  
output:
  pdf_document: default
header-includes: 
 - \usepackage{bm}
 - \usepackage{xcolor}
---

# Portions of Recorded Gibbs Sampler Time for 10 Key Parameters
We first display the first 50 kept post-burn-in MCMC iterations' posterior sampling time (in milliseconds) for 10 key Gibbs sampler steps corresponding to our 4 models, i.e., `fullGPfixedL`, `NNGPblockFixedL`, `NNGPsequenFixedL`, and `NNGPsequenVaryLj`.
```{r echo = -(1:2)}
rm(list=ls())
projDirec <- "C:/Users/annie/OneDrive - National University of Singapore/Documents/PhD/research/first paper/spatempBFA"
wd <- paste(projDirec, "simu/mainScalabilityVerificationSimu/m1600T30K5", sep = "/")
setwd(wd)
load("GibbsStepTimeFixedLfullGP.RData")
load("GibbsStepTimeFixedLblock.RData")
load("GibbsStepTimeFixedLsequen.RData")
load("GibbsStepTimeVaryLjSequen.RData")
head(GibbsStepTimeFixedLfullGP, 50)
head(GibbsStepTimeFixedLblock, 50)
head(GibbsStepTimeFixedLsequen, 50)
head(GibbsStepTimeVaryLjSequen, 50)
```   
As expected, there aren't any significant differences between our 4 models regarding posterior sampling time for the 3 temporal parameters $\psi$, $\Upsilon$, and $\bm{\eta}_t$'s.

# Posterior Sampling Time Summary Statistics
We then present vital posterior sampling time summary statistics for the 7 spatial-related parameters ($z_{jl}(\bm{s}_{i,o})$'s or $u_j(\bm{s}_{i,o})$'s, $\xi_j(\bm{s}_{i,o})$'s, $\theta_{jl_j}$'s, $\delta_{1:k}$, $\rho$, $\kappa$, and $\alpha_{jl_j}(\bm{s}_{i,o})$'s) to showcase the manifest scalability improvements brought about by our 3 novelties, i.e., slice sampling, spatial NNGP, and sequential updates.
```{r}
apply(GibbsStepTimeFixedLfullGP[,1:7], 2, summary)
apply(GibbsStepTimeFixedLblock[,1:7], 2, summary)
apply(GibbsStepTimeFixedLsequen[,1:7], 2, summary)
apply(GibbsStepTimeVaryLjSequen[,1:7], 2, summary)
```
The results correspond well to what we have deduced in Appendix H of our manuscript. 

- Compared to their `fullGPfixedL` counterparts, `NNGPblockFixedL`'s Gibbs sampler steps corresponding to $\rho$ and $\kappa$ are evidently accelerated by our **spatial NNGP prior**;
- The only Gibbs sampler step time that should clearly differ between `NNGPblockFixedL` and `NNGPsequenFixedL` is the step updating all $\alpha_{jl}(\bm{s}_{i,o})$'s, which result from whether we adopt our **sequential updating method** or not. Since $m=1600$ here is big, `NNGPsequenFixedL` is a few times faster than `NNGPblockFixedL` for the posterior sampling step corresponding to $\alpha_{jl}(\bm{s}_{i,o})$'s;
- Thanks to our **slice sampling approach**, `NNGPsequenVaryLj`'s Gibbs sampler steps for $u_j(\bm{s}_{i,o})$'s $\xi_j(\bm{s}_{i,o})$'s, $\rho$, and $\kappa$ are significantly faster than `NNGPsequenFixedL`'s Gibbs sampler steps for $z_{jl}(\bm{s}_{i,o})$'s $\xi_j(\bm{s}_{i,o})$'s, $\rho$, and $\kappa$. It turns out that `NNGPsequenVaryLj`'s Gibbs sampler step for $\alpha_{jl_j}(\bm{s}_{i,o})$'s is slower than its `NNGPsequenFixedL` counterpart, indicating that inefficiencies caused by case discussion, calculating all required upper or lower bounds, and rejection sampling outweigh acceleration brought about by slice sampling's ensured non-increasing posterior samples for $L_j$'s through the MCMC iterations.

We finally calculate standard deviations for the 7 spatial-related parameters' posterior sampling time across all kept post-burn-in MCMC iterations.
```{r}
round(apply(GibbsStepTimeFixedLfullGP[,1:7], 2, sd), 5)
round(apply(GibbsStepTimeFixedLblock[,1:7], 2, sd), 5)
round(apply(GibbsStepTimeFixedLsequen[,1:7], 2, sd), 5)
round(apply(GibbsStepTimeVaryLjSequen[,1:7], 2, sd), 5)
```





     
     


     




